---
title: "assignment1"
output: html_document
---

```{r, echo=FALSE}
setwd("C:/Users/ephyto/Google Drive/coursera - Data Science Specialization/5 Reproducible Research/week 2 assignment/repdata-data-activity")
getwd()
```
```{r setoptions,echo=TRUE}
activity <- read.csv("activity.csv")
dim(activity)
names(activity)
```
###Q1: what is mean total number of steps per day

create new dataset without na values
```{r}
nactivity <- na.omit(activity)
meanTotalSteps <- aggregate(nactivity$steps, list(nactivity$date), sum)
names(meanTotalSteps) <- c("date", "steps")
meanTotalSteps
```
mean line in magenta, median line in blue, median=mean in our example
```{r}
lmts <- nrow(meanTotalSteps)
hist(meanTotalSteps$steps, breaks=lmts)
rug(meanTotalSteps$steps)
abline(v=mean(meanTotalSteps$steps), col="magenta", lwd=4)
abline(v=median(meanTotalSteps$steps), col="blue", lwd=2)
```
```{r}
mean(meanTotalSteps$steps)
```
```{r}
median(meanTotalSteps$steps)
```

###What is the average daily activity
```{r}
library(plyr)
#calculate average steps for each of 5-minute inteval during a 24-hour period
int.mean.steps <- ddply(nactivity, ~interval, summarise, mean=mean(steps))
```
```{r}
library(ggplot2)
qplot(x=interval, y=mean, data = int.mean.steps, geom = "line", xlab="5 minutes inteval", ylab="Number of Step Count", main="Average Number of Steps Across All Days"
)
```
Report 5-min interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
int.mean.steps[which.max(int.mean.steps$mean),]
```
**Observations**

Based on steps taken pattern, the person's daily activity peaks around 8:35am
